---
title: "R_swim_path_plots"
author: "Catarina Pien"
date: "2024-02-08"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, error = FALSE)

# Load packages
library(tidyverse)
library(readxl)
library(boot)
library(gdistance)
library(raster)
library(sf)
library(sp)
library(ggspatial)
library(deltamapr)
library(here)
library(viridis)
library(RColorBrewer)
```


# WY2023

## Load fish data
```{r}
smelt_data_join <- readRDS(here("output/smelt_data_example.rds")) %>%
  mutate(Release = case_when(
                            # Tag == "RBP" ~ "11/30 Soft Release",
                             Tag == "LRA" ~ "11/30 Hard Release",
                             Tag == "ROP" ~ "01/18 Hard Release",
                             Tag == "ad" ~ "01/18 Trailer Release",
                             Tag == "RGP" ~ "01/26 Soft Release",
                             Tag == "LOA" ~ "01/25 Hard Release"),
         Release = factor(Release, levels = c("11/30 Hard Release",
                                              # "11/30 Soft Release",
                                              "01/18 Hard Release",
                                              "01/18 Trailer Release",
                                              "01/25 Hard Release",
                                              "01/26 Soft Release")))

# Read in spatial data
sp_releases <- readRDS(here("output/spatial_release_data_example.rds"))
sp_station <- readRDS(here("output/spatial_station_data_example.rds"))
# raster_baydelta_tr <- readRDS(here("data/raster_baydelta_tr.rds"))
# raster_baydelta<-raster(here("data/baydelta.tif"))
```

## Reorganize data
```{r}
# Subset station to captured delta smelt
sp_station <- sp_station %>% filter(StationCode %in% unique(smelt_data_join$StationCode))

# Make a table of locations A to locations B
sf_release <- st_as_sf(sp_releases) %>%
  mutate(lon1 = sf::st_coordinates(.)[,1],
      lat1 = sf::st_coordinates(.)[,2]) 

sf_catch <- sp_station%>%
  ungroup() %>%
  st_as_sf() %>%
  mutate(lon2 = sf::st_coordinates(.)[,1],
      lat2 = sf::st_coordinates(.)[,2])

releaseToCatch <- smelt_data_join %>%
  left_join(sf_release) %>%
  rename(geometryR = geometry) %>%
  left_join(sf_catch) %>%
  rename(geometryC = geometry)
```

## function for making shortest path lines
```{r}
f_lines <- function(x) {
  paths <- shortestPath(raster_baydelta_tr, c(releaseToCatch$lon1[x], releaseToCatch$lat1[x]), 
                      c(releaseToCatch$lon2[x], releaseToCatch$lat2[x]), output = "SpatialLines")
  line <- sf::st_as_sf(paths)
} 
```

## Run function to make lines - takes a while
```{r, eval = FALSE}
# lineTest <- lapply(1:nrow(releaseToCatch), f_lines)
linesbind <- bind_rows(lineTest) %>% 
  st_as_sf() %>% 
  st_set_crs(st_crs(sp_station)) 
saveRDS(linesbind, file = here("output", "allpaths.rds"), compress = "xz")
```

## Plotting - start here

### Read in data and make df for plotting
```{r}
linesbind <- readRDS(here("output", "allpaths.rds"))
linesbind_df <- cbind(linesbind, 
                      smelt_data_join %>% dplyr::select(SampleDate, Tag, Release.Method, StationCode, FirstDayRelease, Release)) %>%
  mutate(days = as.numeric(SampleDate-FirstDayRelease))

sp_station_Tag <- full_join(sp_station, smelt_data_join %>% dplyr::select(SampleDate, Tag, Release.Method, StationCode, FirstDayRelease, Release))

sp_release_Tag <- full_join(sp_releases, smelt_data_join %>% dplyr::select(SampleDate, Tag, Release.Method, StationCode, FirstDayRelease, Release)) %>%
  filter(Tag %in% unique(smelt_data_join$Tag))
```

### Plotting theme
```{r}
theme_plots <-  
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, size = 10))
```

### Plot all lines
```{r, fig.width = 8, height = 6.5}
plot_allreleases <- ggplot() + 
  geom_sf(data=WW_Delta, fill="gray80", color="gray80") + 
  geom_sf(data=linesbind_df, aes(color = Release)) +
  geom_sf(data=sp_station,color="black", shape = 20)+
  geom_sf(data=sp_releases,color="red", shape = 12)+
  xlim(c(-122.2, -121.3)) +
  ylim(c(37.7, 38.5))+
  annotation_north_arrow(location = "tl", which_north = "true",
                                pad_x = unit(.04, "in"), pad_y = unit(0.05, "in"),
                                style = north_arrow_fancy_orienteering) +
        annotation_scale(location = "bl", bar_cols = c("black", "white", "black", "white")) +
  theme_plots

plot_allreleases
```

### Plot by date
```{r, fig.width = 7, height = 5}
ggplot() + 
  geom_sf(data=WW_Delta, fill="gray80", color="gray80") + 
  geom_sf(data=linesbind_df, aes(color = Release)) +
  geom_sf(data=sp_station_Tag, color="black", shape = 20)+
  geom_sf(data=sp_release_Tag,color="red", shape = 12)+
  xlim(c(-122.2, -121.3)) +
  ylim(c(37.7, 38.5))+
  facet_wrap(~FirstDayRelease) +
  annotation_north_arrow(location = "tl", which_north = "true",
                                pad_x = unit(.04, "in"), pad_y = unit(0.05, "in"),
                                style = north_arrow_fancy_orienteering) +
        annotation_scale(location = "bl", bar_cols = c("black", "white", "black", "white")) +
  theme_plots 
```

### Plot by release and date caught
```{r, fig.width=8, fig.height=6.5}
ggplot() + 
  geom_sf(data=WW_Delta, fill="gray80", color="gray80") + 
  geom_sf(data=linesbind_df, aes(color = SampleDate)) +
  geom_sf(data=sp_station_Tag, color="black", shape = 20)+
  geom_sf(data=sp_release_Tag,color="red", shape = 12)+
  xlim(c(-122.2, -121.3)) +
  ylim(c(37.7, 38.5))+
  facet_wrap(~Release) +
  theme_plots +
  theme(legend.position = "top")
```

### Plot by release and days traveling (similar to above)
```{r, fig.width=8, fig.height=6.5}
ggplot() + 
  geom_sf(data=WW_Delta, fill="gray80", color="gray80") + 
  geom_sf(data=linesbind_df, aes(color = days)) +
  geom_sf(data=sp_station_Tag, color="black", shape = 20)+
  geom_sf(data=sp_release_Tag,color="red", shape = 12)+
  xlim(c(-122.2, -121.3)) +
  ylim(c(37.7, 38.5))+
  facet_wrap(~Release) +
  theme_plots +
  theme(legend.position = "top")
```

### Plot by release and days traveling with viridis
```{r, fig.width=8, fig.height=6.5}
plot_release_days <- ggplot() + 
  geom_sf(data=WW_Delta, fill="gray80", color="gray80") + 
  geom_sf(data=linesbind_df, aes(color = days)) +
  geom_sf(data=sp_station_Tag, color="black", shape = 20)+
  geom_sf(data=sp_release_Tag,color="red", shape = 12)+
  viridis::scale_color_viridis(option = "plasma") +
  xlim(c(-122.2, -121.3)) +
  ylim(c(37.7, 38.5))+
  facet_wrap(~Release) +
  annotation_north_arrow(location = "tl", which_north = "true",
                                pad_x = unit(.02, "in"), pad_y = unit(0.03, "in"),
                                style = north_arrow_fancy_orienteering) +
        annotation_scale(location = "bl", bar_cols = c("black", "white", "black", "white")) +
  theme_plots +
  theme(legend.position = "top")

plot_release_days
```
```{r, eval = FALSE}
png(here("output/Figure_swim_path_release_days.png"), width = 7, height = 6, res = 300, units = "in")
plot_release_days
dev.off()

png(here("output/Figure_swim_path_all.png"), width = 7, height = 6, res = 300, units = "in")
plot_allreleases
dev.off()
```

### Plot for manuscript: Station size based on catch
```{r, fig.width=8, fig.height=6.5}

sp_station_Tag_grouped <- sp_station_Tag %>%
  mutate(Count = 1) %>%
  group_by(Release, StationCode) %>%
  summarize(Catch = sum(Count))


plot_release_days_stationsize <- ggplot() + 
  geom_sf(data=WW_Delta, fill="gray80", color="gray80") + 
  geom_sf(data=linesbind_df, color = "gray30") +
  geom_sf(data=sp_station_Tag_grouped, aes(size = Catch),  color = "magenta4", shape = 10)+
  geom_sf(data=sp_release_Tag,color="blue1", shape = 12)+
  # scale_colour_gradient2(low = "#440154FF",
  #                        mid = "#31688EFF", 
  #                        high = "#6DCD59FF",
  #                        midpoint = 3)+ 
  xlim(c(-122.2, -121.3)) +
  ylim(c(37.7, 38.5))+
  facet_wrap(~Release) +
  scale_size(range = c(1.5, 4)) + 
  annotation_north_arrow(location = "tl", which_north = "true",
                                pad_x = unit(.02, "in"), pad_y = unit(0.07, "in"),
                         height = unit(0.8, "cm"),
                         width = unit(0.8, "cm"),
                                style = north_arrow_fancy_orienteering) +
        annotation_scale(location = "bl", bar_cols = c("black", "white", "black", "white")) +
  theme_plots +
  theme(legend.position = "none")

plot_release_days_stationsize
```

```{r}
png(here("output/Figure_swim_path23.png"), width = 7, height = 6, res = 300, units = "in")
plot_release_days_stationsize
dev.off()
```


# 2024 data 
## Read data
```{r}
data_WY24<-read.csv(here("output", "WY2024_distance_from_release_site.csv"))%>%
  mutate(SampleDate=as.Date(SampleDate)) %>% mutate(DaysinceNov1=as.numeric(SampleDate-as.Date("2022-11-01"))) %>%
  mutate(DaysinceNov1_sqr=DaysinceNov1^2) %>%
  mutate(FirstDayRelease = ymd(FirstDayRelease)) %>%
  mutate(StationCode = if_else(Survey == "TFCF", "CVP", StationCode))
  # mutate(Release = case_when(
  #                           # Tag == "RBP" ~ "11/30 Soft Release",
  #                            Tag == "LRA" ~ "11/30 Hard Release",
  #                            Tag == "ROP" ~ "01/18 Hard Release",
  #                            Tag == "ad" ~ "01/18 Trailer Release",
  #                            Tag == "RGP" ~ "01/26 Soft Release",
  #                            Tag == "LOA" ~ "01/25 Hard Release"),
  #        Release = factor(Release, levels = c("11/30 Hard Release",
  #                                             # "11/30 Soft Release",
  #                                             "01/18 Hard Release",
  #                                             "01/18 Trailer Release",
  #                                             "01/25 Hard Release",
  #                                             "01/26 Soft Release")))

smelt_data_join24 <- data_WY24 %>%
  st_as_sf(coords = c("ReleaseLongitude", "ReleaseLatitude"), crs = 4326, remove = FALSE) %>%
  st_transform(crs = st_crs(sf_release))%>%
  mutate(lon1 = st_coordinates(.)[,1], lat1 = st_coordinates(.)[,2]) %>%
  mutate(Count = 1,
         FirstDayRelease = ymd(FirstDayRelease)) %>%
  dplyr::select(SampleDate, Survey, Count, 
                Tag = `Release.tag.`, 
                Mark = MarkCode, StationCode, FirstDayRelease, 
                `Release Type` = `Release.methods.`, DaySinceRelease, Event, lon1, lat1) %>%
  group_by(SampleDate, Survey, 
                Tag, Mark , StationCode, 
                `Release Type` , FirstDayRelease, DaySinceRelease, Event, lon1 ,lat1) %>%
  summarize(Count = sum(Count)) %>%
  ungroup()

sf_release24 <- data_WY24 %>% 
  mutate(Year = 3) %>%
  dplyr::select(Year, Event, FirstDayRelease, DaySinceRelease, `Release Type` = `Release.methods.`, Tag = `Release.tag.`, lon1 = ReleaseLongitude,lat1 = ReleaseLatitude) %>%
  st_as_sf(coords = c("lon1", "lat1"), crs = 4326) %>%
  st_transform(crs = st_crs(sf_release))%>%
  mutate(lon1 = st_coordinates(.)[,1], lat1 = st_coordinates(.)[,2])

sf_catch24 <- data_WY24%>%
  dplyr::select(StationCode, Survey, Event, lon2=Longitude, lat2=Latitude) %>%
  st_as_sf(coords = c("lon2", "lat2"), crs = 4326)%>%
  st_transform(crs = st_crs(sf_release)) %>%
  mutate(lon2 = st_coordinates(geometry)[,1], lat2 = st_coordinates(geometry)[,2])

sf_station24 <- sf_catch24 %>%
  distinct()

# The row numbers aren't matching up with smelt_data_join24... probably need a sampleID somewhere.
releaseToCatch24 <- smelt_data_join24 %>% st_drop_geometry() %>%
  left_join(sf_release24) %>%
  distinct() %>%
  rename(geometryR = geometry) %>%
  left_join(sf_catch24) %>%
  rename(geometryC = geometry) %>%
  distinct()
```

# Run function to make lines - takes a while
```{r, eval = FALSE}

f_lines24 <- function(x) {
  paths <- shortestPath(raster_baydelta_tr, c(releaseToCatch24$lon1[x], releaseToCatch24$lat1[x]), 
                      c(releaseToCatch24$lon2[x], releaseToCatch24$lat2[x]), output = "SpatialLines")
  line <- sf::st_as_sf(paths)
} 

x=1
result <- shortestPath(raster_baydelta_tr, c(releaseToCatch24$lon1[x], releaseToCatch24$lat1[x]), 
                      c(releaseToCatch24$lon2[x], releaseToCatch24$lat2[x]), output = "SpatialLines")



# lineTest24 <- lapply(1:nrow(releaseToCatch24), f_lines24)
linesbind24 <- bind_rows(lineTest24) %>% 
  st_as_sf() %>% 
  st_set_crs(st_crs(sp_station)) 
saveRDS(linesbind24, file = here("output", "allpaths24.rds"), compress = "xz")
```

# Plotting - start here

## Read in data and make df for plotting
```{r}
linesbind24 <- readRDS(here("output", "allpaths24.rds"))
linesbind24_df <- cbind(linesbind24[-46,], 
                      smelt_data_join24 %>% dplyr::select(SampleDate, Tag, `Release Type`, StationCode, FirstDayRelease, Event)) %>%
  mutate(days = as.numeric(SampleDate-FirstDayRelease))%>%
  mutate(Release = case_when(Event == "LS1" ~ "11/15 LS Hard",
                             Event == "LS2" ~ "1/10 LS Hard",
                             Event == "3a" ~ "1/24 Hard",
                             Event == "3b" ~ "1/25 Soft",
                             Event == "4a" ~ "1/31 Trailer",
                             Event == "4b" ~ "1/30 Soft",
                             Event == "1a" ~ "12/12 Hard",
                             Event == "1b" ~ "12/14 Soft"))%>%
  mutate(Release = factor(Release, levels = c("11/15 LS Hard","12/12 Hard",
                                              "12/14 Soft","1/10 LS Hard","1/24 Hard",
                                              "1/25 Soft","1/30 Soft","1/31 Trailer")))

```

```{r, fig.width=8, fig.height=6.5}
sf_stations_grouped <- sf_catch24 %>%
  mutate(Count = 1) %>%
  group_by(Event, StationCode) %>%
  summarize(Catch = sum(Count)) %>%
  ungroup() %>%
  mutate(Release = case_when(Event == "LS1" ~ "11/15 LS Hard",
                             Event == "LS2" ~ "1/10 LS Hard",
                             Event == "3a" ~ "1/24 Hard",
                             Event == "3b" ~ "1/25 Soft",
                             Event == "4a" ~ "1/31 Trailer",
                             Event == "4b" ~ "1/30 Soft",
                             Event == "1a" ~ "12/12 Hard",
                             Event == "1b" ~ "12/14 Soft"))%>%
  mutate(Release = factor(Release, levels = c("11/15 LS Hard","12/12 Hard",
                                              "12/14 Soft","1/10 LS Hard","1/24 Hard",
                                              "1/25 Soft","1/30 Soft","1/31 Trailer")))
                     

plot_catch24 <- ggplot() + 
  geom_sf(data=WW_Delta, fill="gray80", color="gray80") + 
  geom_sf(data=linesbind24_df, color = "gray30") +
  geom_sf(data=sf_stations_grouped, aes(size = Catch),  color = "magenta4", shape = 10)+
  geom_sf(data=sf_release24,color="blue1", shape = 12)+
  xlim(c(-122.2, -121.3)) +
  ylim(c(37.7, 38.5))+
  facet_wrap(~Release) +
  scale_size(range = c(1.5, 4)) + 
  annotation_north_arrow(location = "tl", which_north = "true",
                                pad_x = unit(.02, "in"), pad_y = unit(0.07, "in"),
                         height = unit(0.8, "cm"),
                         width = unit(0.8, "cm"),
                                style = north_arrow_fancy_orienteering) +
        annotation_scale(location = "bl", bar_cols = c("black", "white", "black", "white")) +
  theme_plots +
  theme(legend.position = "none")

plot_catch24
```

```{r}
png(here("output/Figure_swim_path24.png"), width = 7, height = 8, res = 300, units = "in")
plot_catch24
dev.off()
```
